---
title: "Titanic"
author: "Diego A. Hernandez"
date: "January 28, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## TITANIC

Load train and test sets:
```{r}
setwd("/home/diego/Documents/Kaggle")
train_titanic = read.csv("train.csv")
test_titanic = read.csv("test.csv")
```

Loading libraries:
```{r}
library(MASS)
library(tidyverse)
library(corrplot)
library(forecast)
library(glmnet)
library(FSelector)
library(randomForest)
```

Let's have a quick look of the dataset:
```{r}
glimpse(train_titanic)
```
We clearly see that Survived and Pclass have to be converted into factors:

```{r}
train_titanic$Survived = as.factor(train_titanic$Survived)
train_titanic$Pclass = as.factor(train_titanic$Pclass)

test_titanic$Pclass = as.factor(test_titanic$Pclass)

glimpse(train_titanic)
```

Remove duplicates:

```{r}
length(unique(train_titanic$PassengerId)) == nrow(train_titanic)
```

```{r}
#train_titanic = train_titanic[ , -which(names(train_titanic) %in% c("PassengerId"))]
#test_titanic = test_titanic[ , -which(names(test_titanic) %in% c("PassengerId"))]
```


Counting the number of NAs:
```{r}
na.cols = which(colSums(is.na(train_titanic)) > 0)
sort(colSums(sapply(train_titanic[na.cols], is.na)), decreasing = TRUE)
paste('There are', length(na.cols), 'columns with missing values')
```
We see that Age has 177 missing values.

```{r}
train_titanic %>%
  select(Cabin) %>%
  filter(Cabin !='')

train_titanic = train_titanic[ , -which(names(train_titanic) %in% c("Cabin"))]
test_titanic = test_titanic[ , -which(names(test_titanic) %in% c("Cabin"))]
```
Cabin column has only 204/891 non missing values, we wont play with it. We may delete it.

As we have 177 missing values for Age we are going to build a linear model to predict it:
```{r}
# We divide the train dataset into two: NAs and non-NAs
age_train = train_titanic[!is.na(train_titanic$Age), ]
age_test = train_titanic[is.na(train_titanic$Age), ]

# We use all the variables except Cabin
m_age1 = lm(Age ~ Survived + Pclass + Sex + SibSp + Parch + Fare + Embarked, data = train_titanic) 
summary(m_age1)
```
We see that some of the features are not significants, we will start a backward elimination process to make a good model, lets start by eliminating Sex:
```{r}
m_age2 = lm(Age ~ Survived + Pclass + SibSp + Parch + Fare + Embarked, data = train_titanic)
summary(m_age2)
```
Still non significant, now we will eliminate Embarked:
```{r}
m_age3 = lm(Age ~ Survived + Pclass + SibSp + Parch + Fare, data = train_titanic)
summary(m_age3)
```
Fare is in the limit, more than 5% but still significant at a 10% level of significance, however Parch has a p-value of 24% we will eliminate it:
```{r}
m_age4 = lm(Age ~ Survived + Pclass + SibSp + Fare, data = train_titanic)
summary(m_age4)
```
Right now our model is 100% significant, even Fare.
We will proceed to predict the values and introduce them in the training dataset:
```{r}
age_test$Age = predict(m_age4, newdata = age_test)

train_titanic = age_train %>%
  bind_rows(age_test)

sum(is.na(train_titanic$Age))
```
We have 0 missing values for Age.


After filling the Age column(which is quite relevant), we can start a descriptive analysis of the variables(stats and plots):

```{r}
train_titanic %>%
  group_by(Sex) %>%
  count(Sex)
```
There are 314 females and 577 males, we clearly see a dominant level of males over females in terms of numbers

Lets analyze the Pclass variable and see the dsitribution among richs, middle-class and poors:
```{r}
train_titanic %>%
  group_by(Pclass) %>%
  count(Pclass)
```
As we can see there are more poors passengers than rich, if we want to see it on percentages we can talk about:

Rich = 24.24%
Middle-class = 20.65%
Poor = 55.1%
```{r}
ggplot(data = train_titanic) + geom_bar(aes(x = Pclass, fill = Pclass))
```

Different ports of embarkation:
```{r}
train_titanic %>%
  group_by(Embarked) %>%
  count(Embarked)
```
168 from Cherbourg
77 from Queenstown
644 Southampton
```{r}
ggplot(data = train_titanic) + geom_bar(aes(x = Embarked, fill = Embarked))
```

Survivors:
```{r}
train_titanic %>%
  group_by(Survived) %>%
  count(Survived)
```
More deads than survivors:

Deads = 61,61%
Survivors = 38,39%


Distribution of the ages:
```{r}
ggplot(data = train_titanic, aes(x = Age)) + 
  geom_histogram(aes(colour = I("white"))) + 
  theme_minimal()
```
Range between 20 and 40 is the main characteristic for the Age

**|FEATURE ENGINEERING|**

Let's create a new categorical variable for the Age:

Children = 0-12
Young = 13-28
Middle-age = 29-50
Old > 50

Those values were choosen because back in the 10s life expectancy was quite smaller than nowadays:
```{r}
train_titanic = train_titanic %>%
  mutate(Age_intervals = ifelse(Age >= 0 & Age <= 12, 'Children',
                                ifelse(Age > 12 & Age <= 28, 'Young',
                                       ifelse(Age > 28 & Age <= 50, 'Middle', 'Old'))))
```

After creating this new variable let's analyze it and see how is the distribution:
```{r}
train_titanic %>%
  group_by(Age_intervals) %>%
  count(Age_intervals)
```
Children and Middle-age are majority, this may be linked to the number of families that were on the boat.
```{r}
ggplot(data = train_titanic) + geom_bar(aes(x = Age_intervals, fill = Age_intervals))
```

We are going to create a variable called 'Family' to see if the chances of survival were higher if someone belonged to a family, SibSp == 1 & Parch == 1:
```{r}
train_titanic = train_titanic %>%
  mutate(Family = ifelse(SibSp == 1 & Parch == 1, 1, 0))

test_titanic = test_titanic %>%
  mutate(Family = ifelse(SibSp == 1 & Parch == 1, 1, 0))

glimpse(train_titanic)
#Family and Age_intervals are not factorized let's correct this
train_titanic$Family = as.factor(train_titanic$Family)
train_titanic$Age_intervals = as.factor(train_titanic$Age_intervals)

test_titanic$Family = as.factor(test_titanic$Family)
```
834 Families

**|MODEL|**

We will use the Information Gain information to check if a variable is relevant to our model or not:
```{r}
chi2_titanic = data.frame(chi.squared(formula=Survived ~., data = train_titanic))
chi2_titanic$features = rownames(chi2_titanic)
chi2_titanic[order(chi2_titanic$attr_importance, decreasing = T), ]
```
Name and Ticket are the most important ones but it has no sense cause Ticket it just a code and Name does not influence, let's remove them and re compute the chi2 values:

```{r}
#train_titanic = train_titanic[ , -which(names(train_titanic) %in% c("Name"))]
#train_titanic = train_titanic[ , -which(names(train_titanic) %in% c("Ticket"))]

#test_titanic = train_titanic[ , -which(names(train_titanic) %in% c("Name"))]
#test_titanic = train_titanic[ , -which(names(train_titanic) %in% c("Ticket"))]
```


```{r}
chi2_titanic = data.frame(chi.squared(formula=Survived ~., data = train_titanic))
chi2_titanic$features = rownames(chi2_titanic)
chi2_titanic[order(chi2_titanic$attr_importance, decreasing = T), ]
```

Our variables will be: Sex + Fare + Pclass + Embarked + Age_intervals + Family

```{r}
sampleDF = floor(0.70 * nrow(train_titanic))
set.seed(361)
train_set = sample(seq_len(nrow(train_titanic)), size = sampleDF)

trainTITA = train_titanic[train_set, ]
testTITA = train_titanic[-train_set, ]
```

```{r}
RandomF1 = randomForest(Survived ~ Sex + Fare + Embarked + Age_intervals + Family,
             data = trainTITA)

testing = predict(RandomF1, newdata = testTITA)
#summary(RandomF1)
summary(testing)
```

```{r}
age_train1 = test_titanic[!is.na(test_titanic$Age), ]
age_test1 = test_titanic[is.na(test_titanic$Age), ]

m_age4_test = lm(Age ~ Pclass + SibSp + Fare, data = test_titanic)

age_test1$Age = predict(m_age4_test, newdata = age_test1)

test_titanic = age_train1 %>%
  bind_rows(age_test1)

sum(is.na(test_titanic$Age))

test_titanic = test_titanic %>%
  mutate(Age_intervals = ifelse(Age >= 0 & Age <= 12, 'Children',
                                ifelse(Age > 12 & Age <= 28, 'Young',
                                       ifelse(Age > 28 & Age <= 50, 'Middle', 'Old'))))

test_titanic$Age_intervals = as.factor(test_titanic$Age_intervals)
```

**|FINAL PREDICTION|**
To avoid any problem between training set and test set:
```{r}
levels(test_titanic$Pclass) = levels(train_titanic$Pclass)
levels(test_titanic$Sex) = levels(train_titanic$Sex)
levels(test_titanic$Age_intervals) = levels(train_titanic$Age_intervals)
levels(test_titanic$Embarked) = levels(train_titanic$Embarked)
levels(test_titanic$Family) = levels(train_titanic$Family)
```


Prediction:
```{r}
test_titanic_pred = test_titanic %>%
  mutate(Survived = predict(RandomF1, newdata = test_titanic))
```

Export it as .csv:
```{r}
test_titanic_pred_submission = test_titanic_pred %>%
  select(PassengerId, Survived)

write.csv(test_titanic_pred_submission, "titanic_kaggle.csv")
```


```{r}
test_titanic_pred %>%
  group_by(Survived) %>%
  count(Survived)
```









